
This thesis would not exist without the support of my family, my supervisors and my friends. 
Lesley Gray and Scott Osprey offered me their help, guidance and kindness even before I applied to the DPhil. I am forever grateful for their help and support and that I was able to be part of their group and that I was able to learn from them.
To Robin Chadwick, for always being willing to help me and provide very helpful advice and engage in discussions.

Thanks so much as well to Antje Weishemer, Philip Stier and Tim Woollings for reading my Confirmation and Transfer of Status reports and their helpful feedback which as certainly improved this thesis.  
Many thanks as well for Zane Martin for useful discussions and support throughout this project.
To Andrea and Lucy, for helping me navigate all the administrative processes.

To my parents, Virginia and Jorge, that have always encouraged and supported me all along the way, all together. Thank you. 
Things would be very different without me crossing paths with all the Ambiguos: Karen, Pablo, Diego, Adán, Alinne, Montse, Carolina, Marcos and Dieguito. Thank you for always welcoming me back home. 
Special thanks to Karen Solorzano, Karen Ines and Pablo Bonilla for supporting me at a long distance throughout the difficult times. 
To everyone in Mexico and Leeds that helped me get here: Michel, Angel, Juliane, Paloma, Marcos, Melissa and Ver\'onica. 
 I also wish to thank all the great friends I made in Oxford: Kate, Aaron, RJ, Oscar, Hank, Filippo, Matt and Emily. 

My deepest gratitude goes to the Richard's family, which supported my studies at Oxford through the Oxford-Richards scholarship at Wadham College. Acknowledgements to Wadham College and Dr. Hanna Christensen for all their support and advice. 
To Andrew Bushell, Robin Chadwick and the UK Met Office for the PhD CASE Studentship I was awarded through the Met Office Partnership program which was key to support the science found in this thesis. 




%\subsection*{Institutional}
