%\begin{savequote}[8cm]
%\textlatin{Neque porro quisquam est qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit...}

%There is no one who loves pain itself, who seeks after it and wants to have it, simply because it is pain...
%  \qauthor{--- Cicero's \textit{de Finibus Bonorum et Malorum}}
%\end{savequote}

\chapter{\label{ch:1-intro}Introduction} 

%\minitoc

The American Monsoon System (AMS) is the main source of rainfall for most of
Latin America and the southwestern United States, regions where agriculture is an important economic activity and a vast wealth of ecosystems and biodiversity are
supported by monsoon rainfall in this region. Changes to the amount, timing and location of rainfall over different temporal scales
have direct consequences for society and ecosystems.
For that reason, improving our physical understanding of the mechanisms that cause temporal changes to the AMS rainfall is crucial to improve our medium-range forecasts and our climate predictions which could ultimately render key information for risk assessments, climate adaptation and
agricultural strategies. In this context, this thesis aims to tackle outstanding questions of the AMS using a global climate model with a particular interest in better understanding
the physical mechanisms associated with variability and teleconnections of this monsoon.


\section{Motivation}

Temporal and spatial variability of rainfall is important for society throughout the planet, but any short and long-term changes to precipitation are increasingly relevant in agriculturally active and biodiverse hotspot tropical regions \citep{sultan2005,jain2015}. 
In the AMS, variability on inter-annual scales, i.e., changes to amount or seasonality of rainfall from one year to the next, can produce long-lived droughts that are associated with crop loss and forest fire intensification \citep{chen2009,harvey2018}.
% But shorter-term changes can also produce profound effects such as flash floods and intense rainfall periods that cause crop loss \citep{devereux2007,avila2016recent}.
A large body of monsoon research is consequently focused on improving our understanding of the physical mechanisms responsible for precipitation variability across temporal and spatial scales \citep{wang2017,gadgil2018}. 

The AMS was only recently recognized as a monsoon after the 1990s \added{\citep{adams1997,zhou1998}}, as the definition for a monsoon has evolved from an initial dynamical definition based on the reversal of the prevailing winds to an agronomical definition that recognizes the seasonality of precipitation as the dominant feature of a monsoon \citep{wang2017,gadgil2018}. This  means, however, that our understanding of the AMS is more limited compared to other monsoons given the lower number of studies on the AMS from a monsoon perspective compared to other monsoons such as the Indian monsoons, where monsoon forecasts exist since the 19th century \citep{blanford}.


Recently, theories for monsoon dynamics \citep{bordoni2008monsoons,biasutti2018global,hill2019,geen2020} have arisen to coherently explain the monsoons through a general physical mechanism. Most of these theories aim to explain a global inter-hemispheric band of convection that is driven by the seasonal cycle of solar insolation. Several characteristics of the North and South American monsoons, however, challenge the basic physical inferences or predictions of these theories, which means that these frameworks cannot readily  be applied to the AMS.
\added{For this reason, several primary questions about the AMS remain open such as mechanisms and sources for the seasonal cycle \citep{turrent2009} and intra-seasonal \citep{perdigon2019} and interannual variability \citep{cai2020}, the effect of greenhouse forcing on the monsoon dynamics \citep{liebmann2011,pascale2019} and the role of soil-atmosphere interactions \citep{dominguez2008,malhi2009,adams2013}. The main focus of this thesis is the first topic of the above, i.e., the physical mechanisms for inter-annual and seasonal variability and teleconnections pathways. }% to further understand their variability. 


One prominent example of the lack of understanding of the basic physical mechanisms in the AMS is the puzzling seasonal cycle of rainfall in southern Mexico and Central America. The so-called Midsummer drought is a robust bimodal feature of the seasonal cycle of precipitation during the wet season that has had implications for agricultural practices in the region since the Mayan Empire (AD 800-900) \citep{jobbova2018ritual}. Despite the importance of region-wide agricultural practices, the physical mechanisms that can explain this seasonal variation of rainfall remain disputed over recent years \citep{karnauskas2013,herrera2015,zermeno2019}. 
\added{The topic of seasonal drought mechanisms in southern Mexico and Central America will be a central aspect of this thesis.}

\added{In contrast, a relatively well-known aspect of tropical climate is the El Niño-Southern Oscillation (ENSO), which is a coupled-ocean atmosphere phenomena in the Pacific Ocean that produces long-distance impacts throughout the planet \citep{mcphaden2006}, prominently influencing interannual variability in the AMS regions \citep{magana2003,vera2006,marengo2014climate}.
 Progress has been made in understanding the impacts and pathways of ENSO to South America \citep{marengo2012,cai2020}, and current research has found that diversity in ENSO events causes different impacts to South America \citep{hill2009south,tedeschi2015}. A currently active topic of research aims to understand how impacts to the AMS depend on the interaction between ENSO and its teleconnections  and other aspects of climate variability  \citep{cai2019pantropical,cai2020,jimenez2021}.}

\added{For instance, a growing number of studies have suggested a link between the stratosphere and the tropical troposphere, called stratospheric-tropospheric coupling \citep{haynes2021influence}. This coupling is suggested to include tropospheric features such as ENSO and other modes of tropical variability \citep{hitchman2021observational,martin2021nature}. There is no full description of the causes for these observed relationships yet,  but if the stratosphere is found to be a relevant factor to modulate aspects of tropical convection, such as ENSO impacts, there would be an opportunity to better predict tropospheric variability because the equatorial stratosphere is generally predictable at longer time-scales than the troposphere. Stratospheric-tropospheric coupling will be another central topic of this thesis.}

% Nevertheless, the understanding of the effect of ENSO over South America and the AMS in general is still somewhat limited. For example, two ENSO events that are very similar in the central Pacific can cause teleconnections with different locations and strengths and the reasons behind these varying effects are not well understood. 

% partly because of a lack of long-term reliable data over the Pacific Ocean. 
%the reasons for the observed different responses  to the same long-distance ocean-atmosphere forcing from the Pacific Ocean remain unclear, in spite of the well documented impacts of ENSO which has caused many disastrous floods and droughts across the region.

One key tool to understand the causes for regional changes to monsoon rainfall are general circulation models (GCMs). These models are useful to evaluate the roles of climate features such as orography, air-sea coupling, ENSO teleconnections and impacts, and also monsoons \citep{zhou2016}. However, the use of GCMs to address key questions of the AMS has been scarce and detailed accounts of the biases --differences between the simulated climate of a model and the real world -- are rarely done with explicit emphasis on the AMS. 
In other words, GCMs are rarely evaluated in the AMS, so our understanding is deficient  in the knowledge of the relevant biases in current GCMs and, therefore, GCMs are less frequently used to address scientific questions related to the AMS. 


\added{ This thesis begins with an assessment of the mean state of the AMS in a state-of-the-art climate model in Chapter \ref{ch:4-ams}. This chapter showcases how this model may be used to increase our knowledge of the AMS, by investigating the seasonality of monsoons and bimodal regimes using a new method in Chapter \ref{ch:5-wvt} and the physical mechanisms for the seasonal cycle of southern Mexico in Chapter \ref{ch:6-msd}. The first result chapter also shows suggestive evidence that the stratosphere may play a role in ENSO impacts to South America, which leads to Chapter \ref{ch:7-qbo} which investigates whether there is any coupling between the stratosphere and the troposphere within the GCM. }
% The first results of this thesis begin \added{with Chapter \ref{ch:4-ams}} by evaluating  in the AMS region and comparing the model with several observational datasets. %This , with assessing the roles of biases in the large-scale circulation that affect regional monsoon rainfall biases.
% This assessment highlights that the UM model is fit to investigate two outstanding research questions in the AMS and monsoon literature: the physical mechanisms that control the seasonal cycle of rainfall in southern Mexico and Central America, and the role of the tropical stratosphere for tropical convection and monsoons. 
% 
% For the first research question, Chapter \ref{ch:5-wvt} describes a new method to determine monsoon timings, including the timings of bimodal regimes of precipitation in observations and climate model output. This method is then used in Chapter \ref{ch:6-msd} to investigate the physical mechanisms of the seasonal cycle of precipitation in southern Mexico and Central America.
%  In a way, each chapter of this thesis builds on the previous chapters. 
 % For that purpose, the thesis first presents a chapter that describes the biases and possible processes responsible for these biases in a state-of-the-art climate model. This
 
\section{Thesis aims and outline}
The main aim of this thesis is to investigate the physical causes of variability and the mechanisms associated with teleconnections to the AMS. The specific key aims of this thesis are: 

\begin{enumerate}
\item To characterize the large-scale biases in a state-of-the-art GCM that are relevant for the representation of rainfall in the AMS.  
\begin{enumerate}
\item To characterize the main biases in the thermodynamical and dynamical features over the large scale tropical domain and the regional AMS sub-domains.
\item To evaluate the roles of large-scale biases, horizontal resolution and the use of Earth system processes for regional monsoon representation.
\item To assess the representation of the teleconnection associated with the main driver of interannual variability, i.e., ENSO in a GCM with specific emphasis on the causes for non-linearity and non-asymmetry in the teleconnections.
\end{enumerate}
%To provide a detailed description of the biases in the representation of the AMS in a state-of-the-art climate model to understand 
\item Evaluate the seasonal variability of the monsoon onset, withdrawal and intra-seasonal changes in the GCM and compare to observational datasets. 
\item Describe and investigate the physical mechanisms associated with the seasonal cycle of rainfall in Central America and southern Mexico by testing previous hypothesis of physical mechanisms within the model.
\item To investigate the role of stratospheric-tropospheric coupling in the tropics and the role of the tropopause for convection in the AMS and for ENSO teleconnections.
\end{enumerate}

The remainder of this thesis is structured as follows:

\begin{itemize}
\item Chapter \ref{ch:2-litreview} provides a review of the literature on key aspects of the American monsoons. The chapter begins by introducing the concepts of monsoons, their different physical interpretations as a global phenomena and the place of regional monsoons in the global scale. Then, the North and South American monsoons are introduced and detail is given on the applicability of large-scale monsoon theories to these regional monsoons. This section is followed by a literature review of the proposed physical mechanisms that drive the seasonal cycle of rainfall in Central America, southern Mexico and the Caribbean. El Niño-Southern Oscillation  impacts over North and South America and, finally, the chapter summarises the literature on stratospheric-tropospheric coupling in the tropics, discussing possible mechanisms by which the stratospheric quasi-biennial oscillation may be influential for tropical convection. 

\item  Chapter \ref{ch:3-methods} describes the observational datasets used in this thesis, composed of four gridded precipitation datasets and one reanalysis dataset: ERA5. The chapter also described the UK Met Office Hadley Centre Unified Model (UM) and the configurations of the UM used in this thesis and the Coupled Model Intercomparison Project phase 6 (CMIP6). 

\item Chapter \ref{ch:4-ams} evaluates the representation of the AMS  in three configurations of the UM model submitted to CMIP6. The chapter describes large-scale biases over the tropics and regional scale biases in the precipitation amount and seasonality in key regions of the AMS. 
ENSO teleconnections are also evaluated over the AMS examinging the non-linearity of simulated and observed teleconnections and the role of ENSO flavours. This chapter highlights relevant questions that are of interest to the wider AMS community that are tackled in the remaining chapters: first, the skill of the models in reproducing a bimodal signal in the seasonal cycle of rainfall in Central America and southern Mexico and second, a possible modulation of ENSO teleconnections by the stratospheric quasi-biennial oscillation.
The majority of the work in this chapter has been published in \textit{Weather and Climate Dynamics} as \cite{garciafranco2020}.

\item Chapter \ref{ch:5-wvt} details a wavelet covariant transform method used to diagnose the start and end of the rainy season in a monsoon using precipitation time-series. The method is extended to determine the timing of bimodal regimes, and even whether or not a bimodal regime exists or not in a given region. The method is illustrated in the North American and Indian monsoons, and for the Mid-summer Drought signals of the Caribbean and southern Mexico. This chapter has been published in the \textit{The International Journal of Climatology} as \cite{garciafranco2021}.
\item Chapter \ref{ch:6-msd} uses the wavelet method to investigate the physical mechanisms that cause the bimodal regime of precipitation in southern Mexico in the UM CMIP6 models. The chapter tests elements of three leading hypothesis for the Mid-summer drought, especifically the chapter examines the roles of the East Pacific sea-surface temperatures, the cloud-radiative effects and shortwave radiation and the Caribbean Low-Level Jet. Furthermore, the chapter uses a moist static energy budget, which provides useful insight into the physical mechanisms that domiante intra-seasonal changes to precipitation.
\item Chapter \ref{ch:7-qbo} investigates the tropical route of QBO teleconnections in the pre-industrial control experiments of HadGEM3 and UKESM1 and in targeted nudging experiments.
The observational evidence that links the QBO with the ITCZ, ENSO and the Walker circulation is examined within these simulations to evaluate whether similar relationships to the observations are found within a state-of-the-art GCM. 
The results from these CMIP6 simulations are very similar to observations suggesting robust relationships between the QBO phase, the ITCZs and the Walker circulation.
%\item Chapter \ref{ch:7-nudging} follows up on the topic of QBO teleconnections in the tropics by describing the realization of several experiments using a GCM with a nudged stratosphere.
 The  experiments with a nudged stratosphere improve the biases in the representation of the QBO, so their results can be used to test the causal pathway through which the tropical troposphere and stratosphere interact with each other. 
The results of this chapter show that when the relaxation was applied most of the teleconnections observed in the control experiments disappear, indicating either that the mechanisms that modulate stratospheric-tropospheric coupling were obscured by the experimental design or that the QBO variability is not driving these teleconnections.
\end{itemize}